---
title: "Grouped layouts, rings and flag annotations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Grouped layouts, rings and flag annotations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 7,
  dpi = 96,
  message = FALSE,
  warning = FALSE
)
```

```{r setup}
library(ggvmap)
```

`ggvmap` builds Voronoi *treemaps* — space-filling partitions where each cell's
area is proportional to a weight. Beyond the flat map, three features let you
reproduce publication-style infographics:

1. **Hierarchical (grouped) layouts** — cells organised into sectors by group.
2. **An outer annotation ring** — a labelled coloured band wrapping the map.
3. **Flag / image / value annotations** — pictures and numbers on each cell.

## A flat map

```{r flat}
vm <- voronoi_map(
  weights = c(30, 20, 50, 10, 40),
  labels  = c("Tech", "Health", "Energy", "Finance", "Retail"),
  clip    = clip_circle(),
  seed    = 42
)
autoplot(vm)
```

## Hierarchical layouts and the outer ring

Pass a `group` vector and `voronoi_map()` first divides the disk into one
sector per group (area ∝ the group's total weight), then fills each sector with
its members. `vm_add_ring()` wraps the result in a colour-coded, labelled ring
whose segments line up with the sectors — the world-exports infographic style.

```{r exports}
data(world_exports)

vm <- voronoi_map(
  weights = world_exports$exports,
  labels  = world_exports$country,
  group   = as.character(world_exports$income_group),
  clip    = clip_circle(),
  seed    = 1,
  max_iter = 80
)

autoplot(vm, label_size = 2.3, label_col = "grey20") |>
  vm_add_ring(width = 0.11, label_size = 3.4)
```

The annotation helpers use the `|>` pipe: each reads the underlying map from the
plot and returns a new plot, so they chain naturally.

## Flags and value labels

`vm_add_flags()` resolves country names (English or German) to national flags —
by default via `ggimage::geom_flag()` (`method = "geom_flag"`), or from
[flagcdn.com](https://flagcdn.com) with `method = "url"` (which also supports an
offline `cache = TRUE`). `vm_add_labels()` prints a value
under each name — the merchant-fleet infographic style. (These examples need an
internet connection and the **ggimage** package, so they are not evaluated
here.)

```{r fleet, eval = FALSE}
data(merchant_fleet)

vm <- voronoi_map(
  weights = merchant_fleet$owned,
  labels  = merchant_fleet$country,
  clip    = clip_square(),
  seed    = 3
)

autoplot(vm, label_size = 3.2, label_col = "grey15") |>   # Okabe-Ito by default
  vm_add_labels(
    value = stats::setNames(merchant_fleet$owned, merchant_fleet$country),
    size  = 2.6, nudge_y = -0.035
  ) |>
  vm_add_flags(size = 0.05, nudge_y = 0.05)
```

For offline rendering, pre-download the flags once:

```{r cache, eval = FALSE}
iso   <- country_to_iso(merchant_fleet$country)
files <- flag_cache(iso, dir = "flags")          # writes flags/<iso>.png
```

`vm_add_images()` is the generic version — pass any vector of image paths or
URLs (logos, icons, photos) instead of flags.

## Interactive maps

Set `interactive = TRUE` to make the cells hoverable (highlight + tooltip) via
[ggiraph](https://davidgohel.github.io/ggiraph/), then render the widget with
`vm_girafe()`. It works in R Markdown / Quarto, Shiny and the RStudio viewer.
**Hover the cells below** — this one is live:

```{r interactive, eval = requireNamespace("ggiraph", quietly = TRUE)}
vm <- voronoi_map(
  weights = c(30, 20, 50, 10, 40, 15, 25),
  labels  = c("Tech", "Health", "Energy", "Finance", "Retail", "Media", "Auto"),
  clip    = clip_circle(), seed = 42
)

autoplot(vm, interactive = TRUE) |>
  vm_girafe(width_svg = 6, height_svg = 6)
```

Pass `tooltip =` for custom hover text (length-`n`, named by label, or length 1):

```{r interactive2, eval = FALSE}
autoplot(vm, interactive = TRUE,
         tooltip = paste0(vm$sites$label, ": ", vm$sites$data_weight, "% share")) |>
  vm_girafe()
```
