---
title: "Interactive Voronoi maps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive Voronoi maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  message = FALSE, warning = FALSE,
  fig.width = 6, fig.height = 6
)
run <- requireNamespace("ggiraph", quietly = TRUE)
```

```{r setup}
library(ggvmap)
```

Set `interactive = TRUE` on `autoplot()` or `ggvmap()` to turn any Voronoi map
into an interactive graphic powered by [ggiraph](https://davidgohel.github.io/ggiraph/):
cells **highlight on hover** and show a **tooltip**. Render the result with
`vm_girafe()`. Everything works in R Markdown / Quarto, Shiny and the RStudio
viewer.

## A basic interactive map

Hover the cells — the tooltip shows the label and value.

```{r basic, eval = run}
vm <- voronoi_map(
  weights = c(30, 20, 50, 10, 40, 15, 25),
  labels  = c("Tech", "Health", "Energy", "Finance", "Retail", "Media", "Auto"),
  clip    = clip_circle(), seed = 42
)

autoplot(vm, interactive = TRUE) |>
  vm_girafe(width_svg = 6, height_svg = 6)
```

## Custom tooltips

Pass `tooltip =` (length-`n`, named by label, or length 1) for your own hover
text — HTML is supported.

```{r tooltips, eval = run}
tips <- paste0("<b>", vm$sites$label, "</b><br>", vm$sites$data_weight, "% share")

autoplot(vm, interactive = TRUE, tooltip = tips) |>
  vm_girafe(width_svg = 6, height_svg = 6)
```

## Hierarchical map, interactive

Grouped layouts are interactive too; the default tooltip includes the group.

```{r hier, eval = run}
data(world_exports)

vmh <- voronoi_map(
  weights = world_exports$exports,
  labels  = world_exports$country,
  group   = as.character(world_exports$income_group),
  clip    = clip_circle(), seed = 1, max_iter = 80
)

autoplot(vmh, interactive = TRUE, label_size = 2) |>
  vm_girafe(width_svg = 6.5, height_svg = 6.5)
```

## Notes

* `vm_girafe()` accepts `width_svg` / `height_svg` (inches), a `hover_css`
  string, and extra `ggiraph::opts_*()` via `opts =`.
* The static annotation layers (`vm_add_ring()`, `vm_add_flags()`,
  `vm_add_labels()`) render inside the interactive canvas as well — combine them
  before calling `vm_girafe()`.
* Saved with `htmlwidgets::saveWidget()`, an interactive map is a single
  self-contained `.html` file you can share.

```{r save, eval = FALSE}
g <- autoplot(vm, interactive = TRUE) |> vm_girafe()
htmlwidgets::saveWidget(g, "voronoi.html", selfcontained = TRUE)
```
