---
title: "Accuracy: how close are the areas?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accuracy: how close are the areas?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7,
                      fig.height = 4.2, dpi = 150)
library(ggvmap)
study <- read.csv("fit_study.csv")
pct <- function(x, d = 1) paste0(formatC(100 * x, format = "f", digits = d), "%")
```

A treemap makes one promise: **each cell's area is proportional to its
value**. This article explains how far `ggvmap` keeps that promise, how to
check it for your own map, and what a study of `r nrow(study)` layouts
found.

## Two different questions

**1. Are the cells a proper map, with no gaps and no overlaps?**
Yes, by construction. A point belongs to the site with the smallest
*power distance* $\lVert p - s_i\rVert^2 - w_i$. The border between two
sites, where their power distances are equal, is a straight line, so
each cell is the outline cut by straight lines. Such cells cannot overlap,
leave no gaps and are always convex. Nothing needs to be measured here.
The tests in `tests/testthat/test-correctness.R` still check these
properties on every build, as a guard against programming mistakes.

(This construction is a *power diagram*. It is sometimes confused with
the *additively weighted* Voronoi diagram, which uses
$\lVert p - s_i\rVert - w_i$ and has curved borders.)

**2. Does each cell have the right area?** Only approximately. The weights
$w_i$ are not the data. They are adjusted round by round until the areas
are close to their targets (Nocaj & Brandes 2012, with the step rules of
Franck Lebeau's
[d3-voronoi-map](https://github.com/Kcnarf/d3-voronoi-map)). Iteration
stops when

$$E = \frac{\sum_i |\text{drawn area}_i - \text{target area}_i|}{\text{map area}} < \texttt{convergence\_ratio} \;(= 1\%\text{ by default})$$

or after `max_iter` rounds (200 by default). If the rounds stall (25
rounds without improvement), a second phase holds the sites fixed and
solves for the weights directly with a damped Newton method. For fixed
sites, weights that give exactly the target areas always exist
(Aurenhammer, Hoffmann & Aronov 1998), and this method finds them
reliably (Kitagawa, Mérigot & Thibert 2019).

**This second question is the one that matters**, and it is the subject
of the rest of this article.

## Check your own map

Printing a map reports $E$ and whether it is below the target:

```{r}
vm <- voronoi_map(freshwater$share, labels = freshwater$country,
                  group = freshwater$region, clip = clip_circle(), seed = 5)
vm
```

For grouped maps, half of the allowance goes to dividing the outline among
groups and half to dividing each group among its members. The two errors
can at most add up, so the whole map stays within `convergence_ratio`.
`vm$groups$fit` shows each level:

```{r}
vm$groups$fit
```

`vm_fit()` shows each cell. `target_share` is the share the cell should
cover and `actual_share` the share it does cover. `rel_error` is the
error as a fraction of the cell's own target:

```{r}
fit <- vm_fit(vm)
fit <- fit[order(-abs(fit$rel_error)), ]
head(fit[, c("label", "group", "target_share", "actual_share", "rel_error")])
```

The worst cells are small ones. A cell meant to cover 1% of the map that
covers 1.1% is 10% too big relative to itself, but adds only 0.1% to $E$.
If readers will compare exact values of small cells, check this table or
lower `convergence_ratio`.

One more thing to know: values below `min_weight_ratio` (default 1%) of
the largest value are drawn **at** that floor, so they look bigger than
they are. `target_share` differs from `value_share` exactly for those
cells. Set `min_weight_ratio = 0` to turn this off.

## The study

The script `data-raw/fit_study.R` lays out:

* flat maps of 5, 10, 20 and 40 cells in a square, circle, hexagon and
  triangle;
* flat maps of 100 cells in a circle;
* grouped maps of 30 cells in 5 groups in a circle.

Each setting uses three kinds of values: **even** (Uniform(1, 2), all
within a factor of 2), **typical** (log-normal(3, 1), a few large and many
small) and **extreme** (log-normal(3, 2), spanning several orders of
magnitude). There are 10 random datasets per setting (3 at 100 cells),
and every layout uses the default settings.

```{r}
study$setting <- ifelse(study$type == "grouped", "grouped, 30 cells",
                        paste0("flat, ", study$n, " cells"))
study$setting <- factor(study$setting,
  c(paste0("flat, ", c(5, 10, 20, 40, 100), " cells"), "grouped, 30 cells"))
study$spread <- factor(study$spread, c("even", "typical", "extreme"))
summarise <- function(d) data.frame(
  setting = d$setting[1], values = d$spread[1], layouts = nrow(d),
  converged = sprintf("%d of %d", sum(d$converged), nrow(d)),
  `median rounds` = median(d$iterations),
  `median E` = pct(median(d$total_error), 2),
  `median cell error` = pct(median(d$median_rel_error)),
  `worst cell` = pct(max(d$max_rel_error), 0),
  check.names = FALSE)
out <- do.call(rbind, lapply(split(study, list(study$setting, study$spread),
                                   drop = TRUE), summarise))
knitr::kable(out[order(out$setting, out$values), ], row.names = FALSE)
```

*Converged* is the share of layouts that reached $E < 1\%$.
*Median cell error* is, for each layout, the middle value of
|drawn − target| / target over its cells, then the middle value over
layouts. *Worst cell* is the largest error of any single cell in any
layout of that row.

```{r, fig.alt = "Dot plot of the median cell error of each layout, by setting and value spread."}
library(ggplot2)
ggplot(study, aes(x = 100 * median_rel_error, y = setting, colour = spread)) +
  geom_point(position = position_jitter(height = 0.15, seed = 1),
             alpha = 0.6, size = 1.6) +
  scale_colour_manual(values = c(even = "#1A5B5B", typical = "#F4AB5C",
                                 extreme = "#D1422F")) +
  scale_x_log10() +
  labs(x = "median cell error of one layout (%, log scale)", y = NULL,
       colour = "values",
       title = "How far a typical cell is from its target size",
       subtitle = "one dot per layout") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(), legend.position = "top")
```

**Reading the plot:** each dot is one layout. Its position is how far
that layout's typical cell is from its target size. Dots further right
are less accurate. Colour is the kind of values.

## What we found

```{r, include = FALSE}
worst <- study[which.max(study$max_rel_error), ]
newton_share <- mean(!study$converged_without_newton[study$spread == "extreme"])
big_off <- study$max_rel_error > 0.25
stopifnot(all(study$worst_cell_value_share[big_off] < 1 / study$n[big_off]))
setting_medians <- tapply(study$median_rel_error,
                          interaction(study$setting, study$spread, drop = TRUE),
                          median)
```

* **The target was met in `r sum(study$converged)` of `r nrow(study)`
  layouts.** A layout that misses it reports `Converged: FALSE` when
  printed, so it cannot pass unnoticed.
* **Typical cells are close to their target.** Across settings, the median
  cell error ranged from `r pct(min(setting_medians), 2)` to
  `r pct(max(setting_medians), 1)`. Some layouts with very unequal values
  are especially precise: `r pct(newton_share, 0)` of them were finished by
  the Newton phase, which fits far more tightly than it needs to.
* **Single small cells can be well off.** The worst cell in the whole study
  was drawn at `r formatC(1 + worst$max_rel_error, format = "f", digits = 1)`
  times its target size (a flat map of `r worst$n` cells with
  `r worst$spread` values), in a layout whose total error $E$ was within
  1%. Every cell more than 25% off its target was smaller than the
  average cell of its layout, so such mistakes are hard to see, but they
  matter if exact values are read off small cells. Check `vm_fit()` or lower `convergence_ratio`.
* **The tiling is exact.** In every layout, the cell areas add up to the
  outline's area to within
  `r formatC(max(study$tiling_error), format = "e", digits = 0)`, which is
  rounding error.

## Reproduce it

```r
# from the package's source directory
Rscript data-raw/fit_study.R   # writes vignettes/fit_study.csv
```
