---
title: "ggvmap: honest Voronoi treemaps for ggplot2, in pure R"
author:
  - name: Loukas Theodosiou
    email: loukesio@gmail.com
    corresponding: true
format:
  typst:
    papersize: a4
    fontsize: 10pt
    margin:
      x: 2.4cm
      y: 2.4cm
    number-sections: true
  docx:
    number-sections: true
bibliography: references.bib
reference-section-title: References
execute:
  echo: false
  warning: false
  message: false
abstract: |
  A treemap shows how a whole splits into parts by giving each part an area
  in proportion to its size. A Voronoi treemap does this with irregular,
  many-sided cells inside any convex outline, such as a circle. The cells
  are found by trial and correction, so the areas are never exact, only
  close. ggvmap is an R package that computes these maps in plain R, with
  ggplot2 as its only required dependency, and draws them as ordinary
  ggplot2 plots. Every map reports how close its areas came to the
  intended values, for the map as a whole (`converged`) and for each cell
  (`vm_fit()`). The layout combines the established step rules of the
  d3-voronoi-map library with a second phase, based on Newton's method,
  that rescues maps those rules cannot fit. In a reproducible study of
  519 maps, every map met the default 1% target; without the second phase,
  106 of them would have failed. Next to the two existing R
  implementations, ggvmap is the easiest to install, the fastest for small
  maps, and the only one that reaches its accuracy target at default
  settings. On large maps it is much slower than the compiled
  implementation.
keywords: [data visualization, treemap, Voronoi diagram, power diagram, R, ggplot2]
---

```{r setup}
devtools::load_all("..", quiet = TRUE)
study <- read.csv("../vignettes/fit_study.csv")
bench <- read.csv("figures/benchmark_results.csv")
runs  <- read.csv("figures/benchmark_runs.csv")
pct  <- function(x, d = 1) paste0(formatC(100 * x, format = "f", digits = d), "%")
secs <- function(x) ifelse(x < 1, paste0(round(x * 1000), " ms"),
                    ifelse(x < 90, paste0(round(x, 1), " s"),
                           paste0(round(x / 60, 1), " min")))
flat_small <- study[study$type == "flat" & study$n <= 40, ]
flat_100   <- study[study$type == "flat" & study$n == 100, ]
grouped    <- study[study$type == "grouped", ]
top10 <- freshwater[!grepl("^Rest of|Middle East", freshwater$country), ][1:10, ]
round1 <- voronoi_map(top10$share, clip = clip_circle(), seed = 4, max_iter = 1)
setting_medians <- tapply(study$median_rel_error,
                          paste(study$type, study$n, study$spread), median)
n_failed_before <- sum(!study$converged_without_newton)
failed_spread <- table(factor(study$spread[!study$converged_without_newton],
                              c("even", "typical", "extreme")))
```

# What this package is for

Many questions in science and everyday life are about how a whole splits
into parts: how a country spends its budget, which genes make up a genome,
which countries hold the world's fresh water. A pie chart answers this
with slices. A *treemap* answers it with a picture made of tiles, where
each tile's area is its share of the whole. If Brazil holds 13% of the
world's fresh water, its tile covers 13% of the picture.

The classic treemap uses rectangles. A *Voronoi treemap* [@balzer2005]
uses irregular tiles with many sides instead, and fits them inside any
convex outline: a circle, a hexagon or a triangle (a shape is *convex*
when the straight line between any two of its points stays inside it).
These maps are popular in newspaper graphics because they look organic
and fit any outline. @fig-how shows one with ten countries.

Two R packages already draw Voronoi treemaps: **voronoiTreemap**
[@voronoitreemap], which runs its calculation in a web browser, and
**WeightedTreemaps** [@weightedtreemaps], which runs it in compiled C++
code and needs the CGAL geometry library to be built during
installation. Neither produces a ggplot2 plot, the format most R users
build figures with [@wickham2016]. **ggvmap** fills that gap: the
calculation is written in plain R, ggplot2 is its only required
dependency, and the result is an ordinary ggplot2 object that the rest
of the ggplot2 ecosystem can change.

The rest of this paper explains how the map is built (@sec-how), how close
it comes to the intended areas and how to check that for any map
(@sec-honest), what the package can draw (@sec-use), how it compares with
the other two packages (@sec-bench), and where it falls short
(@sec-limits). The precise mathematics is in @sec-methods.

# How the map is built {#sec-how}

**Pins and pull.** Picture a cork board in the shape of a circle with one
pin per country. Every spot on the board belongs to one pin. In an
ordinary Voronoi diagram, a spot belongs to the nearest pin. Here, each
pin also has a *pull*: a pin with a strong pull claims spots that are a
bit farther away. The exact rule is that a spot belongs to the pin with
the smallest *squared distance minus pull*. Mathematicians call the
resulting picture a *power diagram* [@aurenhammer1987]. Its useful
property is that the border between two neighbouring pins is always a
straight line, so every tile is a convex polygon, and the tiles fit
together with no gaps and no overlaps.

**Adjusting until the areas are right.** At first the pins are spread
out at random and all pulls are similar, so the tiles have the wrong
sizes (@fig-how A). The package then repeats two corrections, one round
at a time [@nocaj2012]:

1. *Move each pin to the middle of its tile.* This keeps tiles round and
   compact rather than long and thin.
2. *Change each pin's pull.* If a tile is smaller than it should be, its
   pull goes up a little; if it is larger, its pull goes down a little.

After each round, the package adds up every tile's mistake (how much too
big or too small it is) and divides by the area of the whole map. When
that total falls below 1% (the default), it stops. If it has not reached
1% after 200 rounds (the default limit), it stops anyway and says so.
@fig-how shows the mistakes shrinking from
`r pct(round1$convergence)`
after the first round to well under 1% at the end.

**Careful steps.** Taking big correction steps makes the tiles overshoot
and wobble back and forth. ggvmap therefore uses the cautious
step rules of Franck Lebeau's JavaScript library *d3-voronoi-map*
[@d3voronoimap], translated into R: a pull never changes by more than
10% per round, steps shrink when the total mistake has gone up in recent
rounds, and a pin whose pull becomes so strong that it would swallow a
close neighbour is corrected. These rules are practical safeguards, not
mathematical guarantees, and they do get stuck: most often when one
value is much larger than all the others, or when the values range from
tiny to huge.

**When the rounds get stuck.** If 25 rounds pass without improvement, the
package stops moving the pins and works out the pulls directly. This
rests on a mathematical result: for pins that stay where they are, there
is always a set of pulls that gives every tile exactly its intended area
[@aurenhammer1998]. Newton's method, the classic technique of improving a
guess by following the slope, finds those pulls reliably if it starts
with all pulls equal and takes careful steps [@kitagawa2019]; in practice
it needs only a handful of steps. Maps that the rounds fit on their own
are not affected.

**Groups.** When values belong to groups (countries in regions, say),
the package first splits the outline into one tile per group, then splits
each group tile among its members with the same method. Each of the two
steps gets half of the 1% allowance, so the finished map as a whole stays
within 1% (@sec-methods explains why).

**Very small values.** A tile for a value hundreds of times smaller than
the largest one would be too small to see or to label. By default, any
value below 1% of the largest is drawn as if it were exactly 1% of the
largest (`min_weight_ratio`). Such tiles are drawn *bigger than their
value*. `vm_fit()` shows which tiles this affected, and
`min_weight_ratio = 0` turns it off.

![How a map is built. The ten countries with the most renewable fresh
water [@faoaquastat], laid out in a circle from the same random start
and stopped after 1, 3 and 6 rounds (A--C) and when finished (D). The
subtitle of each panel gives the area mistakes of all tiles added up, as a
share of the circle; the package stops once this falls below 1%. Pins move
toward the middle of their tiles and pulls are adjusted every round, so
the tiles shift and grow or shrink. (E) For each country (one row), the
open ring is the share of the circle its tile should cover, the yellow dot
is the share it covered after round 1, and the blue dot the share in the
finished map. After round 1 Myanmar's tile is far too small and China's too
large; in the finished map every blue dot sits inside its
ring.](figures/fig0_howitworks.png){#fig-how width="100%"}

# Is the picture honest? {#sec-honest}

A treemap makes one promise: *area equals amount*. Two different
questions decide whether a particular map keeps it.

**Are the tiles a proper map?** Yes, always, because of how they are
built: each tile is cut out of the outline with straight lines, so tiles
cannot overlap, cannot leave gaps, and are always convex. This needs no
experiment. The package's tests still check it on every build
(`tests/testthat/test-correctness.R`), the same way one would check that
a sorting function returns all of its input: as a guard against
programming mistakes, not as a result.

**Does each tile have the right area?** Only approximately, and this is
the question that matters. The rounds stop at a 1% total mistake, or
earlier if they run out of rounds. So every map carries its own report.
Printing the map shows the total mistake and whether the target was met:

```{r echo = TRUE}
vm <- voronoi_map(freshwater$share, labels = freshwater$country,
                  group = freshwater$region, clip = clip_circle(), seed = 5)
vm
```

and `vm_fit()` lists each tile's intended share next to the share it was
drawn with:

```{r echo = TRUE}
fit <- vm_fit(vm)
head(fit[order(-abs(fit$rel_error)), c("label", "value_share",
     "target_share", "actual_share", "rel_error")], 3)
```

The second listing shows the most important lesson about these maps. The
map as a whole is within its 1% allowance, yet single small tiles can be
off by a noticeable *fraction of their own size*: a tile meant to cover
1% of the map that covers 1.1% is 10% too big. A total measure barely
notices this, because the extra area is tiny compared with the whole.
For readers this is usually harmless (no one can see a tenth of a
percent of a circle), but anyone who reads exact values off small tiles
should check `vm_fit()` first, or tighten `convergence_ratio`.

**How close do maps get in practice?** We laid out
`r nrow(study)` maps (@tbl-study): flat maps of 5 to 40 tiles in four
outlines, flat maps of 100 tiles, and grouped maps of 30 tiles in 5
groups, each with three kinds of values: *even* (all within a factor of
two of each other), *typical* (a few large and many small values) and
*extreme* (values spanning several orders of magnitude). Ten different
random datasets were drawn for each setting (three at 100 tiles). The
script is `data-raw/fit_study.R` in the package repository.

```{r}
#| label: tbl-study
#| tbl-cap: "Accuracy study. *Converged*: maps that met the 1% target within the default 200 rounds. *Median tile error*: for each map, the middle value of |drawn − intended| / intended over its tiles; the table gives the middle value over maps. *Worst tile*: the largest such error in any map of the row. Tiles drawn at the `min_weight_ratio` floor are measured against their floored target."
rows <- list(
  "Flat, 5-40 tiles" = flat_small,
  "Flat, 100 tiles"  = flat_100,
  "Grouped, 30 tiles in 5 groups" = grouped)
tab <- do.call(rbind, lapply(names(rows), function(nm) {
  d <- rows[[nm]]
  do.call(rbind, lapply(c("even", "typical", "extreme"), function(sp) {
    e <- d[d$spread == sp, ]
    data.frame(Maps = nm, Values = sp, `Layouts` = nrow(e),
               Converged = sprintf("%d of %d", sum(e$converged), nrow(e)),
               `Median rounds` = median(e$iterations),
               `Median tile error` = pct(median(e$median_rel_error), 2),
               `Worst tile` = pct(max(e$max_rel_error), 0),
               check.names = FALSE)
  }))
}))
tab$Maps[duplicated(tab$Maps)] <- ""
knitr::kable(tab, align = "llrrrrr")
```

Three things stand out. First, the 1% target was met in
`r sum(study$converged)` of `r nrow(study)` maps (@tbl-study,
*Converged*). Without the second phase described in @sec-how, `r n_failed_before`
of these maps would have got stuck (`r failed_spread[["extreme"]]` with
extreme, `r failed_spread[["typical"]]` with typical and
`r failed_spread[["even"]]` with even values), some with area mistakes
adding up to `r pct(max(study$total_error_without_newton), 0)` of the map. Second, the typical tile is drawn very close to its
intended size: the median tile error was
`r pct(median(study$median_rel_error), 2)` over all maps, and below
`r pct(ceiling(100 * max(setting_medians)) / 100, 0)`
in every setting. Third, single small tiles can still be well off. The
worst tile of the study was drawn at
`r formatC(1 + max(study$max_rel_error), format = "f", digits = 1)` times
its intended size, in a map that as a whole was within its 1% allowance
(@tbl-study, *Worst tile*). In every map the tiles covered the outline
exactly (largest difference
`r formatC(max(study$tiling_error), format = "e", digits = 0)`, which is
rounding error in the computer's arithmetic).

**Same values, different picture.** The starting positions of the pins
are random. Different `seed` values therefore give different
arrangements of the same data, all equally valid. The arrangement
carries no meaning: neighbouring tiles are not more similar than
distant ones, and the position of a tile says nothing about its value.
Only areas (and the colours the user chooses) carry information. Setting
`seed` makes a figure reproducible.

# What the package can draw {#sec-use}

**Three steps.** `voronoi_map()` computes the map, `ggvmap()` draws it,
and the `vm_add_*()` functions add decorations: an outer ring naming the
groups (`vm_add_ring()`), extra text such as values (`vm_add_labels()`),
flags (`vm_add_flags()`) and pictures (`vm_add_images()`). Each step
takes the result of the previous one, so they chain with R's pipe
`|>` (@fig-design A--C). Keeping the computation separate from the
drawing is deliberate: the computation is the slow and random part, so it
is done once, checked (@sec-honest), and the same checked map is then
used for every figure made from it.

**Changing the look.** Appearance can be set for the whole map, for one
group, or for one tile (@fig-design D). Where an option takes one value
for everything, it also takes a list of values named after tiles or
groups, so making one country bold, or outlining one region, takes one
argument. Because the result is a ggplot2 plot, titles, themes and fonts
work as for any other ggplot2 figure.

![The design of ggvmap. (A) The steps: values (with optional names and
groups) go into `voronoi_map()`, which computes the map; `ggvmap()` draws
it; decorations are added with the pipe; the result is saved with
`ggsave()` or made interactive with `vm_girafe()`. (B) The nine lines of
code shown produce the finished figure below them from the package's
`freshwater` data (each country's share of the world's renewable fresh
water, 2022 [@faoaquastat]). Each tile is a country, its area is the
country's share, its colour and slice of the circle show the region, and
the outer ring gives each region's total. (C) Every function the package
offers, grouped by purpose; `vm_fit()` is the per-tile check of
@sec-honest, and `vmap_region()` fits maps inside non-convex outlines
such as city boundaries (it needs the optional polyclip package and is
not covered in this paper). (D) Options that change the look, grouped by
whether they act on the whole map, one group or one
tile.](figures/fig1_design.png){#fig-design width="100%"}

@fig-gallery shows the range of maps from one dataset: nine outlines,
grouped maps, the two ring styles, colour by value, and flags.
@fig-usecases shows four examples from different fields. Tiles too small
to hold their name get smaller text (`autoscale`) or no text
(`min_area`), so maps with many small values stay readable without
manual adjustment.

![One dataset, fourteen maps. Rows 1--2: the same ten values in each of
the nine built-in outlines; the outline is a free design choice, and in
each the tile areas match the values. Row 3, left to right: all 30
countries grouped by region, with Latin America outlined in black; a ring
of solid bands naming the regions; a thin ring with each region's total
share; colour showing each country's value, so large values stand out
twice, by area and by colour; and national
flags.](figures/fig2_gallery.png){#fig-gallery width="100%"}

![Example uses, with made-up but realistic numbers; each panel's full
script is in `paper/figures/fig3_usecases.R` and can be reused with real
data. (A) Twelve large gene families, sized by number of genes and grouped
by function. (B) Genetic association hits per human chromosome, in a
rectangle, coloured by the same value as the area. (C) Where the reads of a
sequencing experiment come from, with contamination outlined and the
smallest classes left unlabelled. (D) Household spending, grouped by
need.](figures/fig3_usecases.png){#fig-usecases width="88%"}

# How it compares {#sec-bench}

```{r}
gg <- bench[bench$package == "ggvmap", ]; wt <- bench[bench$package == "WeightedTreemaps", ]
gg <- gg[order(gg$n), ]; wt <- wt[order(wt$n), ]
faster <- gg$n[gg$seconds < wt$seconds]
last_faster <- if (length(faster)) max(faster) else NA
big <- max(gg$n)
gg_conv <- tapply(runs$ggvmap_converged[runs$package == "ggvmap"],
                  runs$n[runs$package == "ggvmap"], function(x) sum(x))
gg_reps <- tapply(runs$rep[runs$package == "ggvmap"],
                  runs$n[runs$package == "ggvmap"], length)
```

@fig-benchmark compares ggvmap with the two other R packages, each at its
default settings, laying out the same values in the same circle.
**Installing** (@fig-benchmark A): ggvmap needs
`r unique(gg$n_deps)` other packages and no compiling; WeightedTreemaps
needs `r unique(wt$n_deps)` plus compiling CGAL; voronoiTreemap needs a
browser environment. **Code** (@fig-benchmark B): the same grouped map
takes about as much code in ggvmap as in WeightedTreemaps, and less than
in voronoiTreemap. **Speed** (@fig-benchmark C): ggvmap is
`r if (!is.na(last_faster)) paste0("faster up to ", last_faster, " tiles (", secs(gg$seconds[gg$n == min(gg$n)]), " versus ", secs(wt$seconds[wt$n == min(wt$n)]), " at ", min(gg$n), " tiles)") else "slower at every size tested"`,
because the compiled package has a fixed start-up cost. It grows slower
quickly with size: at `r big` tiles it took `r secs(gg$seconds[gg$n == big])`
against `r secs(wt$seconds[wt$n == big])`. Treemaps usually show tens of
categories, where the difference is small; for many hundreds of tiles,
WeightedTreemaps is the better tool. **Accuracy** (@fig-benchmark D): by
the total-mistake measure of @sec-how, ggvmap stayed within its 1% target
in `r sum(gg_conv)` of `r sum(gg_reps)` layouts; WeightedTreemaps, at its
own default settings, which stop after a fixed amount of work rather than
at an accuracy target, ended between
`r pct(min(runs$total_error[runs$package == "WeightedTreemaps"]))` and
`r pct(max(runs$total_error[runs$package == "WeightedTreemaps"]))`.
Both packages can be made more accurate by allowing more work, so panel D
compares defaults, not the best each can do.
voronoiTreemap runs inside the browser, so its speed and accuracy cannot
be measured from R.

![Comparison with the two other R implementations, each at its default
settings (protocol in @sec-methods; script
`paper/figures/fig4_benchmark.R`). (A) Other packages that must be
installed first; ggvmap needs only ggplot2 and the packages ggplot2
itself needs. (B) Amount of code for the same grouped map (log scale;
less is better). (C) Time to compute one layout of random values in a
circle, both axes on log scales; faint dots are single datasets, lines
connect the medians. (D) Total area mistake of the finished layout, as in
@fig-how; the dashed line is ggvmap's 1% target. voronoiTreemap appears
only in A--B because it computes in the
browser.](figures/fig4_benchmark.png){#fig-benchmark width="92%"}

# Limitations {#sec-limits}

- **Speed on large maps.** For every tile, ggvmap cuts the outline with
  one straight line per other tile, so the work per round grows with the
  square of the number of tiles. Faster methods are well known: a power
  diagram can be computed from a three-dimensional convex hull in time
  proportional to *n* log *n* [@aurenhammer1987], which is what compiled
  geometry libraries do. We chose the simple method because it fits in a
  few dozen lines of plain R and is fast enough for tens of tiles; the
  price is shown in @fig-benchmark C.
- **No overall guarantee.** The second phase rests on a theorem, but it
  works with whatever pin positions the first phase left, and both
  phases share one budget of rounds (`max_iter`). The package therefore
  always measures the result and says when the target was missed.
- **Small tiles.** A map within 1% overall can still have small tiles
  that are off by a noticeable fraction of their own size, and by default
  tiny values are drawn at a minimum size (@sec-how). Both are visible in
  `vm_fit()`.
- **No labels outside the map.** Tiles too small for text lose their
  label; lines leading to labels outside the map are not yet available.

# Methods {#sec-methods}

**Power diagram.** Each tile $i$ has a site $s_i \in \mathbb{R}^2$ and a
weight $w_i$ (the "pull"); $|x|$ denotes Euclidean length. The power distance from a point $p$ to site
$i$ is $| p - s_i|^2 - w_i$, and tile $i$ is the set of points
of the outline $P$ whose power distance to $i$ is smallest. Setting the
power distances to $i$ and $j$ equal gives
$2(s_j - s_i)\cdot p = | s_j|^2 - | s_i|^2 - w_j + w_i$,
a straight line, so each tile is the intersection of $P$ with $n - 1$
half-planes. ggvmap computes it by clipping $P$ against each half-plane
in turn (Sutherland--Hodgman clipping), $O(n^2)$ work per diagram. This
is a power diagram, not an additively weighted Voronoi diagram, which
uses $| p - s_i| - w_i$ and has curved borders [@balzer2005].
If a tile becomes empty, a tiny placeholder triangle at its site keeps
the bookkeeping intact; its area is far below the target, so such a map
cannot be reported as converged.

**Iteration.** Targets are $A_i = A(P)\,\tilde v_i / \sum_j \tilde v_j$,
where $\tilde v_i = \max(v_i,\ \rho \max_j v_j)$ and $\rho$ is
`min_weight_ratio` (default 0.01). Sites start at well-spread random
positions (best-candidate sampling) and weights at $w_i = A_i/\pi$. Each
round, following @nocaj2012 with the heuristics of @d3voronoimap:
(i) each site moves toward its tile's centroid by a factor
$1 - 0.5f$; (ii) each weight is multiplied by $A_i/a_i$, where $a_i$ is the
current area, clamped to $[0.9 + 0.1f,\ 1.1 - 0.1f]$; (iii) when a
heavier site $h$ and lighter site $l$ satisfy
$| s_h - s_l|^2 < w_h - w_l$, so that $l$'s tile would vanish,
$w_l$ is raised to make the inequality an equality. Here $f$ is the
fraction of the last ten rounds in which the total error increased (0 in
the first ten rounds), which damps steps when the layout oscillates. The
total error is $E = \sum_i |a_i - A_i| / A(P)$; iteration stops when
$E <$ `convergence_ratio` (default 0.01), after 25 rounds without a new
lowest $E$, or after $\texttt{max\_iter} - \min(30, \lfloor\texttt{max\_iter}/4\rfloor)$
rounds (default 200, so 170).

**Second phase.** If the first phase ends without reaching the target,
the sites are fixed at the positions of the round with the lowest $E$ and
the weights are solved for directly. For fixed distinct sites, the map
$w \mapsto a(w)$ is the gradient of a concave function (the dual of a
semi-discrete optimal-transport problem), and weights with $a(w) = A$
exist and are unique up to an additive constant [@aurenhammer1998].
Moving the border between neighbours $i$ and $j$ by changing $w_j$ gives
$\partial a_i / \partial w_j = -L_{ij} / (2 | s_i - s_j |)$,
where $L_{ij}$ is the length of their shared border, and
$\partial a_i/\partial w_i = -\sum_{j \ne i} \partial a_i/\partial w_j$.
Starting from $w = 0$ (an ordinary Voronoi diagram, in which every tile
contains its site and is non-empty), each Newton step solves this system
with $w_1$ fixed and halves the step until every tile keeps at least half
of the smallest starting or target area and the residual
$| A - a |_2$ has fallen by a factor $1 - \tau/2$, where $\tau$
is the step length; this damped Newton method converges globally
[@kitagawa2019]. Phase 2 may use the remaining rounds of `max_iter`.
The returned object records $E$, the number of rounds of both phases,
and `converged` $= (E <$ `convergence_ratio`$)$.

**Grouped maps.** Floored values $\tilde v_i$ are summed per group, the
outline is first divided among groups, and each group tile is then
divided among its members, each level with tolerance
`convergence_ratio`/2. Let $G_g$ and $\hat G_g$ be the target and drawn
area of group $g$, and $\hat A_i = \hat G_g \tilde v_i / \sum_{j \in g}
\tilde v_j$ the member target inside the drawn group tile. By the
triangle inequality,
$\sum_i |a_i - A_i| \le \sum_i |a_i - \hat A_i| + \sum_g |\hat G_g - G_g|$,
so if both levels are within half the tolerance, the whole map is within
the tolerance. `converged` reports the whole-map error; `vm$groups$fit`
reports each level.

**Software tests.** `tests/testthat/test-correctness.R` checks, on random
inputs, that every sampled point lies in a tile whose site has the
smallest power distance, that tiles cover the outline with no gaps or
overlaps, that tiles are convex, and that grouped member tiles fill their
group tile. These are checks of the implementation, since the properties
hold by construction.

**Accuracy study.** `data-raw/fit_study.R` lays out flat maps of
$n \in \{5, 10, 20, 40\}$ in a square, circle, hexagon and triangle;
flat maps of $n = 100$ in a circle; and grouped maps of 30 values in 5
groups in a circle. Values are drawn as Uniform(1, 2) (*even*),
log-normal(3, 1) (*typical*) or log-normal(3, 2) (*extreme*), with 10
datasets per setting (3 at $n = 100$), all at default settings. Results
are in `vignettes/fit_study.csv`.

**Comparison.** Both packages laid out the same log-normal(3, 1) values
in a circle at default settings, for $n$ = 10, 25, 50, 100, 200 (five
datasets each) and 500 (one dataset), timed once per dataset with
`system.time()` on one laptop core. Errors are computed from each
package's own reported tile areas, normalised to shares. Dependency
counts are the recursive Depends, Imports and LinkingTo on CRAN, without
base packages. Code length counts lines, words, characters and function
calls of the shortest code producing the same grouped map. Results:
`paper/figures/benchmark_runs.csv` (per layout) and
`benchmark_results.csv` (medians).

**Availability.** ggvmap is at https://github.com/loukesio/ggvmap under
the MIT licence; the ported d3-voronoi-map heuristics keep their BSD-3
licence notice (`inst/COPYRIGHTS`). Every figure and number in this paper
is produced by the scripts in `paper/figures/` and `data-raw/`, run by
`paper/render.R`.

# Acknowledgements {.unnumbered}

This package grew out of the author's Stack Overflow question ["Make a
circular Voronoi diagram in
R"](https://stackoverflow.com/questions/71799205/make-a-circular-voronoi-diagram-in-r)
and owes much to Allan Cameron's answer there. The iteration heuristics
are Franck Lebeau's, from d3-voronoi-map.

::: {#refs}
:::
